Meta-Analysis of Variation in Sport and Exercise Science

### 2022-08-12 paper/Paper - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Draft to share with everyone

```
# Introduction
Though the quantitative synthesis of results across studies has existed since the 17th century [@plackett_studies_1958], the modern day term "meta-analysis" as it is now referred to was coined by Gene Glass in [-@glass_primary_1976]. Since that time use of meta-analysis as a tool for the synthesis of research in sport and exercise science has increased considerably [@hagger_meta-analysis_2022]. One area of sport and exercise science which has comprised a considerable proportion of that growth appears to have been in the study of resistance training (RT; figure \@ref(fig:trend-plot)). Given such growth, throughout the paper we use RT studies as a hopefully familiar example for sport and exercise science researchers.
```{r trend-plot, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Trends in meta-analyses published in sport and exercise science since 1976"}

@@

## Detecting the presence of interindividual response variation to resistance training intervention
First we conduct a traditional SMD based effect size meta-analysis to explore the effects of RT compared to control for strength outcomes and hypertrophy (i.e., muscle mass/size) outcomes. Polito et al. [-@polito_moderators_2021] originally used a normal random-effects meta-analysis, however the data we extracted were hierarchical in nature (multiple effects within groups within studies) and so a multilevel mixed-effects meta-analysis model with cluster-robust variance estimation was used with random intercepts for study and group[^6]. We then fit the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in in responses differed between RT interventions and controls. A positive SMD would indicate that RT interventions increase outcomes compared to controls, whilst a positive $SD_{ir}$ and $\textrm{ln}VR$ would indicate that the introduction of the RT intervention increased variation in responses compared to controls (i.e., suggests the presence of interindividual response variation).

```{r forest-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of SMD effect sizes for strength (A) and hypertrophy (B) outcomes"}

@@

The pattern of results from our models examining SMDs (figure \@ref(fig:forest-plot)) were similar to those reported by Polito et al. [-@polito_moderators_2021] albeit with slightly lower estimates for both outcome types; this possibly being due to our use of a multilevel mixed-effects meta-analysis model which allowed for each effect size included to be more appropriately weighted. As might be expected, in comparison to non-training controls the RT interventions produced increases in strength (SMD = `r SMD_SDir_logVR$Estimate[1]` [95%CI: `r SMD_SDir_logVR$Lower[1]` to `r SMD_SDir_logVR$Upper[1]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[1]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[1]`) and hypertrophy outcomes (SMD = `r SMD_SDir_logVR$Estimate[4]` [95%CI: `r SMD_SDir_logVR$Lower[4]` to `r SMD_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[4]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[4]`). Confidence intervals were precise for both outcomes, though prediction intervals for SMD estimates (see figure \@ref(fig:forest-plot)) were fairly wide and relative heterogeneity was fairly high coming mostly from between study variance[^7].

In addition to the SMD results, both the $SD_{ir}$ and $\textrm{ln}VR$ were also positive for both strength ($SD_{ir}$ = `r SMD_SDir_logVR$Estimate[2]` [95%CI: `r SMD_SDir_logVR$Lower[2]` to `r SMD_SDir_logVR$Upper[2]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[2]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[2]`; $\textrm{ln}VR$ = `r SMD_SDir_logVR$Estimate[3]` [95%CI: `r SMD_SDir_logVR$Lower[3]` to `r SMD_SDir_logVR$Upper[3]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[3]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[3]`) and hypertrophy outcomes ($SD_{ir}$ = `r SMD_SDir_logVR$Estimate[5]` [95%CI: `r SMD_SDir_logVR$Lower[5]` to `r SMD_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[5]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[5]`; $\textrm{ln}VR$ = `r SMD_SDir_logVR$Estimate[6]` [95%CI: `r SMD_SDir_logVR$Lower[6]` to `r SMD_SDir_logVR$Upper[6]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[6]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[6]`) suggesting that exposure to the RT interventions may have introduced additional variance over and above random error, potentially suggesting the presence of of interindividual response variation. This might support the previous perspectives [@carpinelli_interindividual_2017] that the considerable variation in responses to RT interventions typically observed are due to 'true' interindividual response variation over and above the random error that occurs from pre- and post-intervention measurements (i.e., the variation is *detectable* apart from the random error). However, as noted both the $SD_{ir}$ and $\textrm{ln}VR$ assume constant variance over values of the mean. As we have seen from the SMD analysis RT interventions increase mean scores. Thus, if there is a mean-variance relationship in the data an increase in the mean alone may be fully responsible for any apparent increase in variation. As such, we cannot rely solely on absolute comparisons of variance such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present. The $\textrm{ln}CVR$ can be used to overcome this issue, and below we re-analyse this dataset using this effect size statistic. First though, we present data demonstrating the ubiquity of the mean-variance relationship in typical RT study outcome measures and introduce a model that can also be used to overcome some possible limitations with the $\textrm{ln}CVR$.

@@

```{r mean-variance-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of pre-intervention scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of pre-intervention scores for (C) strength outcomes and (D) hypertrophy outcomes"}

@@

\begin{equation}
\textrm{ln}\hat\sigma_{ij}=(\beta_{0}+\tau_{(1)i}+\tau_{(2)j})+\beta_{1}\textrm{ln}\overline{x}_{ij}+\epsilon_{ij}+m_{ij} 
(\#eq:lnm-lns-model-eq)
\end{equation}

where $\textrm{ln}\hat\sigma_{ij}$ is the effect size, as in equation \@ref(eq:lnsigma-eq), from the $j\textrm{th}$ group ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of groups) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\textrm{ln}\overline{x}_{ij}$ is the mean estimate for each effect size, $\beta_{0}$ is the intercept, $\beta_{1}$ is the slope or regression coefficient for $\textrm{ln}\overline{x}$, $\tau_{i}$ is the deviation from $\beta_{0}$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ group, and $\epsilon_{ij}$ is the residual for each effect size which is normally distributed with $\sigma^2_{\epsilon}$, and $m_{ij}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}\hat\sigma_{ij}}$. Additional predictor terms could be added to this model; for example, we could model a categorical variable for the outcome type and include $(\beta_{2} + \varphi_{i} + \varphi_{j})Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome groups), and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ group. In this case $\tau_{i}$ and $\varphi_{i}$, and $\tau_{j}$ and $\varphi_{j}$ are assumed to have multivariate normal distributions with the following variance-covariance structure:

@@

```{r model-mean-variance-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of pre-intervention scores"}

@@

## Reanalysis of interindividual response variation using $lnCVR$ and the random slope mixed effects model
As can be seen in figure \@ref(fig:mean-variance-delta-plot) there is a mean-variance relationship in the change score data whereby an increase in the mean alone (i.e., greater mean change score in the intervention compared to the control) may be fully responsible for any apparent increase in variation. As such, we cannot rely solely on absolute comparisons of variance such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present.

```{r mean-variance-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of change scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of change scores for (C) strength outcomes and (D) hypertrophy outcomes"}

@@

The $\textrm{ln}CVR$ can be used to overcome this issue though. Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation was used with random intercepts for study and group and before using the $\textrm{ln}CVR$ as the response variable leads to different conclusions compared to absolute variance comparisons. The introduction of an RT intervention actually *reduces* the relative variation seen in change scores for strength ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[28]` [95%CI: `r mods_SMD_logCVR$Lower[28]` to `r mods_SMD_logCVR$Upper[28]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[28]`, $I^2_{group}$ = `r mods_SMD_logCVR$I2.group[28]`) and hypertrophy ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[88]` [95%CI: `r mods_SMD_logCVR$Lower[88]` to `r mods_SMD_logCVR$Upper[88]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[88]`, $I^2_{group}$ = `r mods_SMD_logCVR$I2.group[88]`).

@@

\begin{equation}
\textrm{ln}CVR_{ij}=\mu+\tau_{i}+\tau_{j}+m_{ij} 
(\#eq:lnCVR-model-eq)
\end{equation}

where $\textrm{ln}CVR_{ij}$ is each effect size, as in equation \@ref(eq:lnCVR-eq), for the $j\textrm{th}$ group ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of groups) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\mu$ is the intercept or overall mean, $\tau_{i}$ is the deviation from $\mu$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ group, which are assumed to be normall distributed around zero with variance of $\sigma^2_{\tau}$, and $m_{ij}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}CVR_{ij}}$. Instead, we can use a version of the random slope model described above (equation \@ref(eq:lnm-lns-model-eq) and following paragraph) to compare the variability between intervention and control groups using $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$. In this case, the categorical variable for the outcome type is instead swapped for the group and the new model term included becomes $(\beta_{2} + \varphi_{i} + \varphi_{j})Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$, and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ group.

@@

Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)) we fit this model to the dataset. The results were however largely similar to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_log_mean_mod_strength$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_log_mean_mod_strength$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$ci.ub[3],2)`]). See figure \@ref(fig:model-mean-variance-delta-plot).

```{r model-mean-variance-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=8, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of change scores"}

@@

The reason for this apparent reduction in variation after introduction of an RT intervention is not necessarily discernible from this analysis. Perhaps the introduction of an RT intervention has indirect effects that reduce other sources of random variance (e.g., diet, other physical activity etc.; [@halliday_resistance_2017]), or a ceiling effect on change (i.e., plateau in response; [@steele_long-term_2022]) has a constraining effect [@cortes_martinez_constant_2021]. Hopefully it is clear from the regression model here where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) that there is considerable flexibility in the inclusion of predictors (i.e., meta-regression) when exploring variance through a meta-analytic framework. Models can be fit to explore not only how study and group level characteristics moderate effect size estimates when considering not only SMDs, but also when considering the variance based effect size statistics and models employed in this article[^8].

@@

@@

[^6]: Data was coded such that study and group had explicit nesting.

[^7]: Note, we also explored for signs of small study bias, including publication bias favouring the finding of intervention effects, for the SMDs given that the relative lack of awareness for variance based effect sizes in the field implies that they might have more influence over such biases. There did not appear to be any obvious small study bias in the dataset (ADD LINK TO SUPPLEMENTARY)

[^8]: See supplementary materials (ADD LINK TO SUPPLEMENTARY) for model estimates for both SMD and $\textrm{ln}CVR$ (used as results for $\textrm{ln}CVR$ and the random slope model were similar) across a range of categorical and continous predictors for both strength and hypertrophy outcomes. There were no obvious moderators of $\textrm{ln}CVR$ in particular.

>>>>

```
# Introduction
Though the quantitative synthesis of results across studies has existed since the 17th century [@plackett_studies_1958], the modern day term "meta-analysis" as it is now referred to was coined by Gene Glass in [-@glass_primary_1976]. Since that time use of meta-analysis as a tool for the synthesis of research in sport and exercise science has increased considerably [@hagger_meta-analysis_2022]. One area of sport and exercise science which has comprised a considerable proportion of that growth appears to have been in the study of resistance training (RT; figure \@ref(fig:trend-plot)). Given such growth, throughout the paper we use RT studies as a hopefully familiar example for sport and exercise science researchers.
```{r trend-plot, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Trends in meta-analyses published in sport and exercise science since 1976."}

@@

## Detecting the presence of interindividual response variation to resistance training intervention
First we conduct a traditional SMD based effect size meta-analysis to explore the effects of RT compared to control for strength outcomes and hypertrophy (i.e., muscle mass/size) outcomes[^6]. Polito et al. [-@polito_moderators_2021] originally used a normal random-effects meta-analysis, however the data we extracted were hierarchical in nature (multiple effects within groups within studies) and so a multilevel mixed-effects meta-analysis model with cluster-robust variance estimation was used with random intercepts for study and group[^7]. We then fit the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in in responses differed between RT interventions and controls. A positive SMD would indicate that RT interventions increase outcomes compared to controls, whilst a positive $SD_{ir}$ and $\textrm{ln}VR$ would indicate that the introduction of the RT intervention increased variation in responses compared to controls (i.e., suggests the presence of interindividual response variation).

```{r forest-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of SMD effect sizes for strength (A) and hypertrophy (B) outcomes."}

@@

The pattern of results from our models examining SMDs (figure \@ref(fig:forest-plot)) were similar to those reported by Polito et al. [-@polito_moderators_2021] albeit with slightly lower estimates for both outcome types; this possibly being due to our use of a multilevel mixed-effects meta-analysis model which allowed for each effect size included to be more appropriately weighted. As might be expected, in comparison to non-training controls the RT interventions produced increases in strength (SMD = `r SMD_SDir_logVR$Estimate[1]` [95%CI: `r SMD_SDir_logVR$Lower[1]` to `r SMD_SDir_logVR$Upper[1]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[1]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[1]`) and hypertrophy outcomes (SMD = `r SMD_SDir_logVR$Estimate[4]` [95%CI: `r SMD_SDir_logVR$Lower[4]` to `r SMD_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[4]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[4]`). Confidence intervals were precise for both outcomes, though prediction intervals for SMD estimates (see figure \@ref(fig:forest-plot)) were fairly wide and relative heterogeneity was fairly high coming mostly from between study variance.

In addition to the SMD results, both the $SD_{ir}$ and $\textrm{ln}VR$ were also positive for both strength ($SD_{ir}$ = `r SMD_SDir_logVR$Estimate[2]` [95%CI: `r SMD_SDir_logVR$Lower[2]` to `r SMD_SDir_logVR$Upper[2]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[2]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[2]`; $\textrm{ln}VR$ = `r SMD_SDir_logVR$Estimate[3]` [95%CI: `r SMD_SDir_logVR$Lower[3]` to `r SMD_SDir_logVR$Upper[3]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[3]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[3]`) and hypertrophy outcomes ($SD_{ir}$ = `r SMD_SDir_logVR$Estimate[5]` [95%CI: `r SMD_SDir_logVR$Lower[5]` to `r SMD_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[5]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[5]`; $\textrm{ln}VR$ = `r SMD_SDir_logVR$Estimate[6]` [95%CI: `r SMD_SDir_logVR$Lower[6]` to `r SMD_SDir_logVR$Upper[6]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[6]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[6]`) indicating that exposure to the RT interventions may have introduced additional variance over and above random error, potentially suggesting the presence of interindividual response variation. This might support previous perspectives [@carpinelli_interindividual_2017] that the considerable variation in responses to RT interventions typically observed are due to 'true' interindividual response variation over and above the random error that occurs from pre- and post-intervention measurements (i.e., the variation is *detectable* apart from the random error). However, as noted both the $SD_{ir}$ and $\textrm{ln}VR$ assume constant variance over values of the mean. As we have seen from the SMD analysis RT interventions increase mean scores. Thus, if there is a mean-variance relationship in the data an increase in the mean alone may be fully responsible for any apparent increase in variation. As such, we cannot rely solely on absolute comparisons of variance such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present. The $\textrm{ln}CVR$ can be used to overcome this issue, and below we re-analyse this dataset using this effect size statistic. First though, we present data demonstrating the ubiquity of the mean-variance relationship in typical RT study outcome measures and introduce a model that can also be used to overcome some possible limitations with the $\textrm{ln}CVR$.

@@

```{r mean-variance-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of pre-intervention scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of pre-intervention scores for (C) strength outcomes and (D) hypertrophy outcomes."}

@@

\begin{equation}
\textrm{ln}\hat\sigma_{ijk}=(\beta_{0}+\tau_{(1)i}+\tau_{(2)j})+\beta_{1}\textrm{ln}\overline{x}_{ijk}+\epsilon_{ijk}+m_{ijk} 
(\#eq:lnm-lns-model-eq)
\end{equation}

where $\textrm{ln}\hat\sigma_{ij}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnsigma-eq), from the $j\textrm{th}$ group ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of groups) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\textrm{ln}\overline{x}_{ijk}$ is the mean estimate for each effect size, $\beta_{0}$ is the intercept, $\beta_{1}$ is the slope or regression coefficient for $\textrm{ln}\overline{x}$, $\tau_{i}$ is the deviation from $\beta_{0}$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ group, and $\epsilon_{ijk}$ is the residual for each effect size which is normally distributed with $\sigma^2_{\epsilon}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}\hat\sigma_{ijk}}$. Additional predictor terms could be added to this model; for example, we could model a categorical variable for the outcome type and include $(\beta_{2} + \varphi_{i} + \varphi_{j})Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome groups), and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ group. In this case $\tau_{i}$ and $\varphi_{i}$, and $\tau_{j}$ and $\varphi_{j}$ are assumed to have multivariate normal distributions with the following variance-covariance structure:

@@

```{r model-mean-variance-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of pre-intervention scores."}

@@

## Reanalysis of interindividual response variation using $\textrm{ln}CVR$ and the random slope mixed effects model
As can be seen in figure \@ref(fig:mean-variance-delta-plot) there is a mean-variance relationship in the change score data whereby an increase in the mean alone (i.e., greater mean change score in the intervention compared to the control) may be fully responsible for any apparent increase in variation. As such, we cannot rely solely on absolute comparisons of variance such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present.

```{r mean-variance-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of change scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of change scores for (C) strength outcomes and (D) hypertrophy outcomes."}

@@

The $\textrm{ln}CVR$ can be used to overcome this issue though. Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation and random intercepts for study and group as before using the $\textrm{ln}CVR$ as the effect size statistic leads to different conclusions compared to absolute variance comparisons. The introduction of an RT intervention actually *reduces* the relative variation seen in change scores for strength ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[28]` [95%CI: `r mods_SMD_logCVR$Lower[28]` to `r mods_SMD_logCVR$Upper[28]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[28]`, $I^2_{group}$ = `r mods_SMD_logCVR$I2.group[28]`) and hypertrophy ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[88]` [95%CI: `r mods_SMD_logCVR$Lower[88]` to `r mods_SMD_logCVR$Upper[88]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[88]`, $I^2_{group}$ = `r mods_SMD_logCVR$I2.group[88]`) and further there is lower relative heterogeneity between studies in the effect estimates.

@@

\begin{equation}
\textrm{ln}CVR_{ijk}=\mu+\tau_{i}+\tau_{j}+m_{ij} 
(\#eq:lnCVR-model-eq)
\end{equation}

where $\textrm{ln}CVR_{ij}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnCVR-eq), for the $j\textrm{th}$ group ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of groups) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\mu$ is the intercept or overall mean, $\tau_{i}$ is the deviation from $\mu$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ group, which are assumed to be normall distributed around zero with variance of $\sigma^2_{\tau}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}CVR_{ijk}}$. Instead, we can use a version of the random slope model described above (equation \@ref(eq:lnm-lns-model-eq) and following paragraph) to compare the variability between intervention and control groups using $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$. In this case, the categorical variable for the outcome type is instead swapped for the group and the new model term included becomes $(\beta_{2} + \varphi_{i} + \varphi_{j})Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$, and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ group.

@@

Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)) we fit this model to the dataset. The results were largely similar, albeit slightly attenuated, to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_log_mean_mod_strength$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_log_mean_mod_strength$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$ci.ub[3],2)`]). See figure \@ref(fig:model-mean-variance-delta-plot).

```{r model-mean-variance-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=8, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of change scores."}

@@

Hopefully it is clear from the regression model here where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) that there is considerable flexibility in the inclusion of predictors (i.e., meta-regression) when exploring variance through a meta-analytic framework. Models can be fit to explore not only how study and group level characteristics moderate effect size estimates when considering not only SMDs, but also when considering the variance based effect size statistics and models employed in this article[^8].

@@

The reason for this apparent reduction in variation after introduction of an RT intervention is not necessarily discernible from this analysis. Perhaps the introduction of an RT intervention has indirect effects that reduce other sources of random variance (e.g., diet, other physical activity etc.; @halliday_resistance_2017), or a ceiling effect on change (i.e., plateau in response; @steele_long-term_2022) has a constraining effect [@cortes_martinez_constant_2021]. This potentially represents another interesting area of future study regarding variation however; how to produce interventions that actually reduce variance in an outcome. In other contexts such as sporting performance, interventions to not only positively affect mean performance but also those that reduce variation in performance would be highly desirable.

@@

[^6]: Note, we also explored for signs of small study bias, including publication bias favouring the finding of intervention effects, for the SMDs given that the relative lack of awareness for variance based effect sizes in the field implies that they might have more influence over such biases. There did not appear to be any obvious small study bias in the dataset (see https://osf.io/stqr3).

[^7]: Data was coded such that study and group had explicit nesting.

[^8]: See supplementary materials (https://osf.io/e6vpr) for model estimates for both SMD and $\textrm{ln}CVR$ (used as results for $\textrm{ln}CVR$ and the random slope model were similar) across a range of categorical and continous predictors for both strength and hypertrophy outcomes. There were no obvious moderators of $\textrm{ln}CVR$ in particular.
